---
title: "HW04 Part 2: Complete Chapter 3"
author: "Jordan Beardslee"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r}
library("tidyverse")
```

### Plot <1.>
```{r}
?trees
```

```{r}
ggplot(data = trees) +
  geom_point(mapping = aes(x = Height, y = Girth))
```

###Plot <2.>
```{r}
?year
```

```{r}
year <- c(1821:1934)
```

```{r}
class(lynx)
as.vector(year)
```
```{r}
pelts <- c(as.vector(year))
```
```{r}
lynx_pelts <- data.frame(year, pelts)
```
```{r}
ggplot(data = lynx_pelts, mapping = aes(x = year, y = pelts, color = "maroon")) + geom_line() +
  geom_smooth(se = FALSE)
```

###Plot <3.>
```{r}
as.vector(iris)
```


```{r}
ggplot(data = iris) +
  geom_point(mapping = aes(x = Petal.Length, y = Petal.Width, color = Species, shape = Species, size = 2)) +
  labs(x = "Petal Length", y = "Petal Width")
```

###Plots <4.> <5.>
```{r}
?geom_violin
```

```{r}
ggplot(data = msleep) +
  geom_violin(fill = "grey70", mapping = aes(x = vore, y = sleep_rem))
```
This plot shows where more of the data is clustered by the wider portion of the violin shapes. The wider the poriton, the more data is clustered in that area.

###Plot <7.>
```{r}
ggplot(data = msleep, mapping = aes(x = order, y = sleep_total)) + 
  geom_boxplot() + coord_flip()
  
```

###Plot <8.>
```{r}
ggplot(data = msleep, mapping = aes(x = conservation, y = awake)) + 
  geom_boxplot() + geom_point() + geom_jitter() + coord_flip()+
  labs(x = "Conservation", y = "Awake")
```

###Plots <9.> and <10.>

```{r}
ggplot(data = msleep) + 
  geom_point(mapping = aes(x = sleep_total, y = brainwt, color = sleep_cycle)) + facet_wrap(~ sleep_rem, nrow = 2)
```
By the looks of this graph brain weight does not have much to do with sleep total. Most of the data points are clustered around 10 hours, and vary in both weight and sleep cycle differences.

```{r}
ggplot(data = msleep) + 
  geom_point(mapping = aes(x = sleep_total, y = bodywt, color = sleep_rem)) + facet_wrap(~ sleep_cycle, nrow = 2)
```
Increased body weight appears to have some affect on total sleep time with higher weights spending less total time asleep and in rem sleep. as weights decrease total time asleep and in rem apprears to increase. 

