---
title: "HW05 Part 1: Complete the sections"
author: "Jordan Beardslee"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

- Change "your name" in the YAML header above to your name.

- As usual, enter the examples in code chunks and run them, unless told otherwise.

## Chapter 10: Tibbles

Read [R4ds Chapter 10: Tibbles](https://r4ds.had.co.nz/tibbles.html), sections 1-3.

### 10.1: Introduction

Load the tidyverse package. 
```{r}
library(tidyverse)
```


### 10.2: Creating tibbles

Enter your code chunks for Section 10.2 here.

Describe what each chunk code does. 
```{r}
as_tibble(iris)
```
as_tibble(iris) created a small table by category.

```{r}
tibble(x = 1:5, y = 1, z = x ^ 2 + y)
```
this function narrowed down the table to just the three categories I wanted. 
```{r}
tb <- tibble(
  `:)` = "smile", 
  ` ` = "space",
  `2000` = "number"
)
tb
```
This created a table with categories named the way i wanted, instead of how R wanted. 

```{r}
tribble(
  ~x, ~y, ~z,
  #--|--|----
  "a", 2, 3.6,
  "b", 1, 8.5
)
```


### 10.3: Tibbles vs data.frame

Enter your code chunks for Section 10.2 here.

Describe what each chunk code does. 

```{r}
tibble(
  a = lubridate::now() + runif(1e3) * 86400,
  b = lubridate::today() + runif(1e3) * 30,
  c = 1:1e3,
  d = runif(1e3),
  e = sample(letters, 1e3, replace = TRUE))
```
This shows only the columns that will fit on the screen and only the first ten rows. This makes it easier to find the data I would want without hunting through a massive table.

```{r}
nycflights13::flights %>% 
  print(n = 10, width = Inf)
```
This condenses the flights into a table that is not only easy to read but fits on the screen.
```{r}
nycflights13::flights %>% 
  View()
```
Viewing the flight table shows that it is very large, using the print function the table becomes more manageable.
```{r}
df <- tibble(
  x = runif(5),
  y = rnorm(5)
)

df$x

df[["x"]]

df[[1]]
```
This allows me to pull a single variable from a data frame instead of having to hunt for it.

```{r}
df %>% .$x

df %>% .[["x"]]
```


### 10.4: Not required

#### Section 10.5 Questions

Answer the questions *completely.* Use code chunks, text, or both, as necessary.

**1:** How can you tell if an object is a tibble? (Hint: try printing `mtcars`, which is a regular data frame). Identify at least two ways to tell if an object is a tibble. *Hint:* What does `as_tibble()` do? What does `class()` do? What does `str()` do?
```{r}
as_tibble(mtcars)
```
When we put in mtcars it gives all the collums, but using as_tibble gives us the first ten instead of everything. It also gives only what can fit on the screen.


**2:** Compare and contrast the following operations on a data.frame and equivalent tibble. What is different? Why might the default data frame behaviours cause you frustration?

```{r}
df <- data.frame(abc = 1, xyz = "a")
df$x
df[, "xyz"]
df[, c("abc", "xyz")]
```
```{r}
tbl <- as_tibble(df)
tbl$x
tbl[, "xyz"]
tbl[, c("abc", "xyz")]
```
The dataframe appears to know exactly what to do with the df$x but the tibble seems thrown off by it. My guess would be that a data frame can take what data there is from that variable, while the tibble can't.

## Chapter 11: Importing data

Read [R4ds Chapter 11: Data Import](https://r4ds.had.co.nz/data-import.html), sections 1, 2, and 5.

### 11.1 Introduction

Nothing to do here unless you took a break and need to reload `tidyverse`.

### 11.2 Getting started.

Do *not* run the first code chunk of this section, which begins with `heights <- read_csv("data/heights.csv")`. You do not have that data file so the code will not run.

Enter and run the remaining chunks in this section.

```{r}
read_csv("a,b,c
1,2,3
4,5,6")
```
```{r}
read_csv("The first line of metadata
  The second line of metadata
  x,y,z
  1,2,3", skip = 2)

read_csv("# A comment I want to skip
  x,y,z
  1,2,3", comment = "#")
#> # A tibble: 1 x 3
#>       x     y     z
#>   <dbl> <dbl> <dbl>
#> 1     1     2     3
```
```{r}
read_csv("1,2,3\n4,5,6", col_names = FALSE)
#> # A tibble: 2 x 3
#>      X1    X2    X3
#>   <dbl> <dbl> <dbl>
#> 1     1     2     3
#> 2     4     5     6
```
```{r}
read_csv("1,2,3\n4,5,6", col_names = c("x", "y", "z"))
#> # A tibble: 2 x 3
#>       x     y     z
#>   <dbl> <dbl> <dbl>
#> 1     1     2     3
#> 2     4     5     6
```

```{r}
read_csv("a,b,c\n1,2,.", na = ".")
#> # A tibble: 1 x 3
#>       a     b c    
#>   <dbl> <dbl> <lgl>
#> 1     1     2 NA
```


#### 11.2 Questions

**1:** What function would you use to read a file where fields were separated with "|"?

You would use read_delim("|")

**2:** (This question is modified from the text.) Finish the two lines of `read_delim` code so that the first one would read a comma-separated file and the second would read a tab-separated file. You only need to worry about the delimiter. Do not worry about other arguments. Replace the dots in each line with the rest of your code. 


# Comma-separated
`file <- read_csv("file.csv", ...)`

# Tab-separated
`file <- read_tsv("file.csv", ...)`


**3:** What are the two most important arguments to `read_fwf()`? Why?
Col_positions is the most important argument because it tells where the data column begins and ends.

**4:** Skip this question


**5: ** Identify what is wrong with each of the following inline CSV files. What happens when you run the code?

```{r}
read_csv("a,b\n1,2,3\n4,5,6")
read_csv("a,b,c\n1,2\n1,2,3,4")
read_csv("a,b\n\"1")
read_csv("a,b\n1,2\na,b")
read_csv("a;b\n1;3")
```

1.) Only header "a" and "b" are designated as columns, but thre are three rows of data.

2.) The number of columns in the data does not match the number of columns in the header.

3.) There are more headers than there are data columns.

4.)"a" and "b" are treated as numbers in the data as well as headers.

5.)The values are seperated by semi colons but read_csv2() isn't being used.

### 11.3 and 11.4: Not required

### 11.5: Writing to a file

Just read this section. You may find it helpful in the future to save a data file to your hard drive. It is basically the same format as reading a file, except that you must specify the data object to save, in addition to the path and file name.

### 11.6 Not required

## Chapter 18: Pipes

Read [R4ds Chapter 18: Pipes](https://r4ds.had.co.nz/pipes.html), sections 1-3.

Nothing to do otherwise for this chapter. Is this easy or what?

**Note:** Trying using pipes for all of the remaining examples. That will help you understand them.

## Chapter 12: Tidy Data

Read [R4ds Chapter 12: Tidy Data](https://r4ds.had.co.nz/tidy-data.html), sections 1-3, 7. 

### 12.1 Introduction

Nothing to do here unless you took a break and need to reload the `tidyverse.`

### 12.2 Tidy data

Study Figure 12.1 and relate the diagram to the three rules listed just above them. Relate that back to the example I gave you in the notes. Bear this in mind as you make data tidy in the second part of this assignment.

You do not have to run any of the examples in this section.

### 12.3

Read and run the examples through section 12.3.1 (gathering), including the example with `left_join()`. We'll cover joins later.

#### 12.3 Questions

**2:** Why does this code fail? Fix it so it works.

```{r}
table4a %>% 
  gather(`1999`, `2000`, key = "year", value = "cases")
#> Error in inds_combine(.vars, ind_list): Position must be between 0 and n
```
The years did not have the ` mark to designate them as variables. The code was not recognizing them as nubmers.

That is all for Chapter 12. On to the last chapter.


## Chapter 5: Data transformation

Read [R4ds Chapter 5: Data Transformation](https://r4ds.had.co.nz/transform.html), sections 1-4.

Time to [get small.](https://www.youtube.com/watch?v=GOrdzCHnpw4) 

### 5.1: Introduction

Load the necessary libraries. As usual, type the examples into and run the code chunks.
```{r}
library(nycflights13)
library(tidyverse)
```

```{r}
flights
```
This is a very condensed version of the actual table, using only the first few rows of data, and the first few columns.

### 5.2: Filter rows with `filter()`

Study Figure 5.1 carefully. Once you learn the `&`, `|`, and `!` logic, you will find them to be very powerful tools.

```{r}
filter(flights, month == 1, day == 1)
```
This gave me only the flights on the first day of the first month.

```{r}
jan1 <- filter(flights, month ==1, day == 1)
```

```{r}
(dec25 <- filter(flights, month == 12, day == 25))
```
This gives me only the flights on the 25th of December, filtering out all the other days.

```{r}
near(sqrt(2) ^ 2,  2)
#> [1] TRUE
near(1 / 49 * 49, 1)
#> [1] TRUE
```
This is simply giving me the closest number that the computer can in order to keep from using infinite numbers.

```{r}
filter(flights, month == 11 | month == 12)
```
This code uses the "|" to decide that the flights between November and December will be included in the data.

```{r}
nov_dec <- filter(flights, month %in% c(11, 12))
```

```{r}
filter(flights, !(arr_delay > 120 | dep_delay > 120))
filter(flights, arr_delay <= 120, dep_delay <= 120)
```
This line of code filtered out all the flights that were delayed, leaving only those that arrived on time to be included in the data. 

#### 5.2 Questions

**1.1:** Find all flights with a delay of 2 hours or more.
```{r}
filter(flights, (arr_delay > 120 | dep_delay > 120))
```


**1.2:** Flew to Houston (IAH or HOU)
```{r}
filter(flights, dest == 'IAH' | dest == 'HOU')
```


**1.3:** Were operated by United (UA), American (AA), or Delta (DL).
```{r}
filter(flights, carrier == 'UA' | carrier == 'AA' | carrier == 'DL')
```


**1.4:** Departed in summer (July, August, and September).
```{r}
filter(flights, month >= 7 & month <= 9)
```


**1.5:** Arrived more than two hours late, but didn’t leave late.

```{r}
filter(flights, arr_delay > 120, dep_delay <= 0)
```


**1.6:** Were delayed by at least an hour, but made up over 30 minutes in flight. This is a tricky one. Do your best.

```{r}
filter(flights, dep_delay >= 60, dep_delay-arr_delay > 30)
```

**1.7:** Departed between midnight and 6am (inclusive)
```{r}
filter(flights, dep_time <=600 | dep_time == 2400)
```


**2:** Another useful dplyr filtering helper is `between()`. What does it do? Can you use it to simplify the code needed to answer the previous challenges?
```{r}
filter(flights, between(month, 7, 9))
filter(flights, !between(dep_time, 601, 2359))
```


**3:** How many flights have a missing dep_time? What other variables are missing? What might these rows represent?
```{r}
summary(flights)
```
8255 have missing dep times, 8255 have missing dep_delay, 8713 have missing arr_time, 9430 have missing air_time

**4:** Why is `NA ^ 0` not missing? Why is `NA | TRUE` not missing? Why is `FALSE & NA` not missing? Can you figure out the general rule? (`NA * 0` is a tricky counterexample!)

NA ^ 0 is equal to 1. NA | TRUE means that if either term is true, the whole set is counted as true. FALSE & NA is not missing because one of the answers in the expression is still false. The general rule seems to be if at least part of the set is correct, it is counted as a correct piece of code. 

**Note:** For some context, see [this thread](https://blog.revolutionanalytics.com/2016/07/understanding-na-in-r.html)


### 5.3 Arrange with `arrange()`


#### 5.3 Questions

**1:** How could you use `arrange()` to sort all missing values to the start? (Hint: use is.na()). **Note:** This one should still have the earliest departure dates after the `NA`s. *Hint:* What does `desc()` do?

```{r}
df <- tibble(x = c(5, 2, NA))
arrange(df, desc(is.na(x)))
arrange(df, -(is.na(x)))
```


**2:** Sort flights to find the most delayed flights. Find the flights that left earliest. 

This question is asking for the flights that were most delayed (left latest after scheduled departure time) and least delayed (left ahead of scheduled time).

```{r}
arrange(flights, desc(dep_delay))
arrange(flights, dep_delay)

```



**3:** Sort flights to find the fastest flights. Interpret fastest to mean shortest time in the air.
```{r}
flights %>% mutate(travel_time = ifelse((arr_time - dep_time < 0), 2400+(arr_time - dep_time), arr_time - dep_time)) %>% arrange(travel_time) %>% select(arr_time, dep_time, travel_time)
```


*Optional challenge:* fastest flight could refer to fastest air speed. Speed is measured in miles per hour but time is minutes. Arrange the data by fastest air speed.


**4:** Which flights travelled the longest? Which travelled the shortest?

```{r}
arrange(flights, desc(distance)) %>% select(1:5, distance)
arrange(flights, distance) %>% select(1:5, distance)
```

### 5.4 Select columns with `select()`

#### 5.4 Questions

**1:** Brainstorm as many ways as possible to select `dep_time`, `dep_delay`, `arr_time`, and `arr_delay` from flights. Find at least three ways.
```{r}
select(flights, dep_time,  dep_delay, arr_time, arr_delay)
select(flights, c(dep_time,  dep_delay, arr_time, arr_delay))
flights %>% select(dep_time,  dep_delay, arr_time, arr_delay)
```


**2:** What happens if you include the name of a variable multiple times in a `select()` call?

Nothing since its the same variable.


**3:** What does the `one_of()` function do? Why might it be helpful in conjunction with this vector?
It brings back all the variables you used. This would be helpful so you know exactly which variables you have used in the vector.

`vars <- c("year", "month", "day", "dep_delay", "arr_delay")`

flights %>% select(one_of(vars))


**4:** Does the result of running the following code surprise you? How do the select helpers deal with case by default? How can you change that default?

`select(flights, contains("TIME"))`

```{r}
select(flights, contains("TIME"))


```
The results aren't that surprising since everything in the table contains the word time, the helpers just pick everything that has the word time in it. 

