---
title: "HW 08: Dates"
author: "Jordan Beardslee"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r}
library(tidyverse)

library(lubridate)
library(nycflights13)
```
####Reading ch16 sections 1-3
```{r}
today()

now()
```
```{r}
ymd("2019-01-04")

mdy("January 4th, 2019")

dmy("04-Jan-2019")
```
```{r}
ymd(20190104)
```

```{r}
ymd_hms("2019-01-04 20:52:00 CST")

mdy_hm("01/04/2019 08:53")
```
```{r}
ymd(20170121, tz = "UTC")
```

```{r}
flights %>%
  select(year, month, day, hour, minute)
```

```{r}
flights %>%
  select(year, month, day, hour, minute) %>%
  mutate(departure = make_datetime(year, month, day, hour, minute))
```
```{r}
make_datetime_100 <- function(year, month, day, time) {
  make_datetime(year, month, day, time %/% 100, time %% 100)
}

flights_dt <- flights %>% 
  filter(!is.na(dep_time), !is.na(arr_time)) %>% 
  mutate(
    dep_time = make_datetime_100(year, month, day, dep_time),
    arr_time = make_datetime_100(year, month, day, arr_time),
    sched_dep_time = make_datetime_100(year, month, day, sched_dep_time),
    sched_arr_time = make_datetime_100(year, month, day, sched_arr_time)
  ) %>% 
  select(origin, dest, ends_with("delay"), ends_with("time"))

flights_dt
```
```{r}
flights_dt %>%
  ggplot(aes(dep_time)) +
  geom_freqpoly(binwidth = 86400)
```
```{r}
flights_dt %>% 
  filter(dep_time < ymd(20130102)) %>% 
  ggplot(aes(dep_time)) + 
  geom_freqpoly(binwidth = 600) # 600 s = 10 minutes
```
```{r}
as_datetime(today())

as_date(now())
```
```{r}
as_datetime(60 * 60 * 10)

as_date(365 * 10 + 2)
```

```{r}
datetime <- ymd_hms("2016-07-08 12:34:56")

year(datetime)

month(datetime)

mday(datetime)

yday(datetime)

wday(datetime)
```
```{r}
month(datetime, label = TRUE)

wday(datetime, label = TRUE, abbr = FALSE)
```
```{r}
flights_dt %>% 
  mutate(wday = wday(dep_time, label = TRUE)) %>% 
  ggplot(aes(x = wday)) +
    geom_bar()
```
```{r}
flights_dt %>% 
  mutate(minute = minute(dep_time)) %>% 
  group_by(minute) %>% 
  summarise(
    avg_delay = mean(arr_delay, na.rm = TRUE),
    n = n()) %>% 
  ggplot(aes(minute, avg_delay)) +
    geom_line()
```
```{r}
sched_dep <- flights_dt %>% 
  mutate(minute = minute(sched_dep_time)) %>% 
  group_by(minute) %>% 
  summarise(
    avg_delay = mean(arr_delay, na.rm = TRUE),
    n = n())

ggplot(sched_dep, aes(minute, avg_delay)) +
  geom_line()
```
```{r}
ggplot(sched_dep, aes(minute, n)) +
  geom_line()
```
```{r}
flights_dt %>% 
  count(week = floor_date(dep_time, "week")) %>% 
  ggplot(aes(week, n)) +
    geom_line()
```

```{r}
(datetime <- ymd_hms("2016-07-08 12:34:56"))

year(datetime) <- 2020
datetime

month(datetime) <- 01
datetime

hour(datetime) <- hour(datetime) + 1
datetime
```
```{r}
update(datetime, year = 2020, month = 2, mday = 2, hour = 2)
```

```{r}
ymd("2015-02-01") %>% 
  update(mday = 30)

ymd("2015-02-01") %>% 
  update(hour = 400)
```

```{r}
flights_dt %>% 
  mutate(dep_hour = update(dep_time, yday = 1)) %>% 
  ggplot(aes(dep_hour)) +
    geom_freqpoly(binwidth = 300)
```

####Part 1
```{r}
birth_date <- "11 March 1992"

# Elizabeth Garret Anderson was the first female physician
anderson <- "9 June 1836"

# Mary Anning discovered the first complete Plesiosaur fossil, 
# and other important Jurassic fossils
anning <-  "1799/05/21"

# Alice Augusta Ball was a chemist who developed the
# first injectable treatment for leprosy.
ball <- "July 24, 1892"

# Sylvia Earle was the first chieft scientist at NOAA
earle <- "August, 30 (1935)"

# Rosalind Franklin used x-ray diffraction to 
# discover DNA structure. Watson and Crick claimed credit.
franklin <- "25th of July, 1920"

# Ester Lederberg was the first to discover that viruses 
# infect bacteria. Led to discovery of lambda bacteriophage
lederberg <- "December 18, 1922"

# Barbara McClintock discovered transposable elements (jumping genes),
# and developed staining techniques to study chromosomes
mcclintock <- "16061902"

# Martin Luther King's birthday is coming up. Let's remember a dream
mlk_birth <- "1/15/29"
mlk_dream <- "1963 August 08"
```
#1. Convert to POSIXct format using lubridate functions.
```{r}
beardslee <- dmy(birth_date)
anderson <- dmy(anderson)
anning <- ymd(anning)
ball <- mdy(ball)
earle <- mdy(earle)
franklin <- dmy(franklin)
lederberg <- mdy(lederberg)
mcclintock <- dmy(mcclintock)
mlk_birth <- mdy(mlk_birth)
mlk_dream <- ymd(mlk_dream)
mlk_birth <- update(mlk_birth, year = 1929)

beardslee_dttm <- as_datetime(beardslee)
anderson_dttm <- as_datetime(anderson)
anning_dttm <- as_datetime(anning)
ball_dttm <- as_datetime(ball)
earle_dttm <- as_datetime(earle)
franklin_dttm <- as_datetime(franklin)
lederberg_dttm <- as_datetime(lederberg)
mcclintock_dttm <- as_datetime(mcclintock)
mlk_birth_dttm <- as_datetime(mlk_birth)
mlk_dream_dttm <- as_datetime(mlk_dream)

beardslee_dttm
anderson_dttm
anning_dttm
ball_dttm
earle_dttm
franklin_dttm
lederberg_dttm
mcclintock_dttm
mlk_birth_dttm
mlk_dream_dttm
```
#2 Check the dates to make sure they are correct.  
Dates were correct but mlk_birth needed updating to fix year in order to show up correctly. 
#3 What day of the week was I born on?
```{r}
wday(beardslee, label = TRUE, abbr = FALSE)
```
I was born on a Wednesday.

#4 What day of the week will my birthday be this year?
```{r}
what_year <- year(today())
update(beardslee, year = what_year)
wday(beardslee, label = TRUE, abbr = FALSE)
```
Looks like it falls on another Wednesday.

#5 What day of the week will my birthday fall on in 2525?
```{r}
beardslee %>%
  update(year = 2525) %>%
  wday(label = TRUE, abbr = FALSE)
```
My birthday will be on a Sunday.

#How many days between the birth of Anderson and Ball
```{r}
anderson - ball
```
-20499 days between Anderson and Ball.

#7 Duration between MLK birth and MLK dream
```{r}
Between <- mlk_dream - mlk_birth
Between <- as.duration(Between)
Between
```
34.56 years approximately. 

#Enter the eight birthdates into vectors
```{r}
birth_vector <- c(beardslee_dttm, anderson_dttm, anning_dttm, ball_dttm, earle_dttm, franklin_dttm, lederberg_dttm, mcclintock_dttm, mlk_birth_dttm)
names(birth_vector) <- c("Beardslee", "Anderson", "Anning", "Ball", "Earle", "Franklin", "Lederberg", "McClintock", "King")
```

#Which dates were leap years?
```{r}
leapyears <- leap_year(birth_vector)
leapyears
year(birth_vector[leapyears])
```
There were three leap years in the data with Franklin (1920), Anderson (1836), and Ball(1892) being born on leap years. There were 4 all together since my own birthday (1992) fell on a leap year.

#Subtract each birthdate in the vector from your own to calculate the number of days that have passed. 
```{r}
days_passed_1 <- time_length(birth_vector - beardslee_dttm, "day")
sort(days_passed_1, decreasing = FALSE)

```

#Part 2
```{r}
library(readr)
library(dplyr)
library(tidyverse)

lake_ice
```

###Filter the data to save only Lake Mendota
```{r}
lake_mendota <- filter(lake_ice, lakename == "LAKE MENDOTA")
lake_mendota
```

#Select the columns
```{r}
select(lake_mendota, iceon_year, iceon_month, iceon_day, iceoff_year, iceoff_month, iceoff_day)
```
```{r}
lake_mendota_1 <- lake_mendota %>% 
  select(iceon_year, iceon_month, iceon_day, iceoff_year, iceoff_month, iceoff_day) %>% 
  mutate(ice_on = make_date(iceon_year, iceon_month, iceon_day),
         ice_off = make_date(iceoff_year, iceoff_month, iceoff_day))
lake_mendota_1
```

#Create a new duration column by subtracting ice
```{r}
lake_mendota_2 <- lake_mendota1 %>% 
  mutate(duration = as.numeric(ice_off - ice_on))
lake_mendota_2
```

#Line Graph
```{r}
ggplot(data = lake_mendota_2) +
  geom_line(mapping = aes(x = iceoff_year, y = duration), na.rm = TRUE)+
  coord_cartesian(xlim = c(1850, 2030))
```

#Soil Carbon Dioxide

```{r}
soil_carbon <- read.delim("~/bi485/beardslee_jordan/data/soil_co2.txt")
soil_carbon
```

#Establish date variable
```{r}
jan <- make_date(year = 2006, day = 1, month = 1)
```

#Rename the columns
```{r}
soil_levels <- soil_carbon %>%
  rename(Days = DayN..) %>% 
  rename(Dateco = Datecode..) %>% 
  rename(Min = Min..) %>% 
  rename(Veg = Veg..) %>% 
  rename(Plot = Plot..) %>% 
  rename(Group = Group..) %>% 
  rename(Concentration = CorrCO2..) %>% 
  rename(Stype = SoilT..)
```

#Date column
```{r}
soil_date <- soil_levels %>%
   mutate(Date = jan + Days,
         Year = year(Date),
         Month = month(Date, label = TRUE, abbr = TRUE)) %>% 
  group_by(Year, Month)

soil_date
```

#Summarize
```{r}
soil_sum <- soil_date %>%
  summarise(
    meanCO2 = mean(Concentration),
    Co2se = sd(Concentration)/sqrt(NROW(soil_date))
  )
soil_sum
```

#Column chart
```{r}
ggplot(data = soil_sum) +
  geom_col(aes(x = Month, y = meanCO2)) +
  geom_linerange(aes(x = Month, ymin = meanCO2 - Co2se, ymax = meanCO2 + Co2se)) +
  facet_wrap(~Year, nrow = 2)
```

#Column with points
```{r}
ggplot(data = soil_sum) +
  geom_point(aes(x = Month, y = meanCO2)) +
  geom_linerange(aes(x = Month, ymin = meanCO2 - Co2se, ymax = meanCO2 + Co2se)) +
  facet_wrap(~Year, nrow = 2)

```

#Mississippi River Discharge

```{r}
thebes <- read_csv("~/bi485/beardslee_jordan/data/thebes_discharge.csv")

thebes

colnames(thebes)
thebes_col <- thebes %>%
  gather ("1988", "1989", "1990", "1991", "1992", "1993", "1994", "1995", "1996", "1997", "1998", "1999", "2000", "2001", "2002", "2003","2004", "2005", "2006", "2007", "2008", "2009", "2010", "2011", "2012", "2013", "2014", "2015", "2016", "2017", "2018", key = "Year", value = "discharge") %>% 
  group_by(Year, Month) %>% 
  mutate(discharge = discharge / 1000,
         Event = case_when(
           Year== "1993" ~ "Flood",
           Year == "2011" ~ "Flood",
           Year == "1988" ~ "Drought",
           Year == "2012" ~ "Drought",
           TRUE ~ "Normal"),
         average_discharge = mean(discharge), na.rm = FALSE)
thebes_col
```


#Plot 2
```{r}
events <- thebes_col %>%
  filter(Event %in% c("Flood", "Drought"))
events
```

#Plot the data
```{r}
plot_1 <- thebes_col %>%
  filter(!Year %in% c("1988", "1993", "2011", "2012")) %>%
  ggplot() +
  geom_line(aes(x = Month, y = average_discharge, group = Year),
            color = "gray70",
            size = 0.3) +
  labs(x = NULL,
       y = "Discharge (cf/s, thousands)")
plot_1
```
# Add second layer
```{r}
thebes_col %>%
  filter(!Year %in% c("1988", "1993", "2011", "2012")) %>%
  ggplot() +
  geom_line(aes(x = Month, y = average_discharge, group = Year),
            color = "gray70",
            size = 0.3) +
  labs(x = NULL,
       y = "Discharge (cf/s, thousands)") +
  geom_line(data = events, aes(x = Month, y = average_discharge, group = Year, color = Event))
```


