---
title: "HW 07 Part 1"
author: "Jordan Beardslee"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
chunk_output_type: inline
---

```{r}
library(tidyverse)
library(GGally)
library(devtools)

```

```{r}
devtools::install_github("thomasp85/patchwork")
```
####Part1: Recreate the Zuu figures
```{r}
library(patchwork)
```
```{r}
sparrow_raw <- read_tsv("~/bi485/beardslee_jordan/data/sparrows.txt")

sparrows <- sparrow_raw %>% 
  mutate(box_group = "Sparrows",
         row_order = 1:nrow(.))

as.tibble(sparrows)
```
####Figure 2: outliers
```{r}
plot_box <- sparrows %>%
  ggplot(aes(x = box_group, 
             y = wingcrd)) +
  geom_boxplot(width = 0.25) +
  labs(x = NULL,
       y = "Wing length (mm)")
```

```{r}
plot_cleveland <- sparrows %>%
  ggplot(aes(x = wingcrd,
             y = row_order)) +
  geom_point() +
  labs(x = "Wing length (mm)",
       y = "order of data")
  
```

```{r}
plot_box + plot_cleveland
```
```{r}
plot_box + plot_cleveland + coord_flip()
```

####Figure 3: outliers
```{r}
p_culmen <- sparrows %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = culmen)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
p_culmen
```

```{r}
p_nalospi <- sparrows %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = nalospi)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
p_nalospi

```


```{r}
p_wt <- sparrows %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = wt)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
p_wt

```

```{r}
p_wingcrd <- sparrows %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = wingcrd)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
p_wingcrd

```

```{r}
p_tarsus <- sparrows %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = tarsus)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
p_tarsus

```

```{r}
p_head <- sparrows %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = head)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
p_head

```

```{r}
p_culmen + p_nalospi + p_wt + p_wingcrd + p_tarsus + p_head + plot_layout(ncol = 3)
```
Question: Head, Nalospi, and tarsus appear to have outlier data.

####Figure 5: distribution
```{r}
h1 <- sparrows %>%
  filter(Month %in% 6:8) %>% 
  ggplot() +
  geom_histogram(aes(x = wt),
                 binwidth = 0.5,
                 boundary = 0,
                 closed = "right",
                 color = "black") +
  scale_x_continuous(breaks = seq(14, 28, by = 2))
```

```{r}
h2 <- sparrows %>%
  filter(Month %in% 6:8) %>% 
  mutate(Month = case_when(
    Month == 6 ~ "June",
    Month == 7 ~ "July",
    Month == 8 ~ "August")) %>% 
  ggplot() +
  geom_histogram(aes(x = wt),
                 binwidth = 0.5,
                 boundary = 0,
                 color = "black") +
  scale_x_continuous(breaks = seq(14, 28, by = 2)) +
  facet_wrap(~ Month, ncol = 1)
```

```{r}
h1 + h2
```
####Figure 10: Relationships

```{r}
sparrows %>% ggscatmat(columns = 1:7)
```
```{r}
month_order = c("May", "June", "July", "August", "September")
sparrows %>% 
  filter(Month %in% 5:9,
         Sex != 0) %>% 
  mutate(Month = case_when(
    Month == 5 ~ "May",
    Month == 6 ~ "June",
    Month == 7 ~ "July",
    Month == 8 ~ "August",
    Month == 9 ~ "September"),
    Month = factor(Month, levels = month_order, ordered = TRUE),
    Sex = ifelse(Sex == 4, "Male", "Female")) %>% 
  ggplot(aes(x = wingcrd,
             y = wt)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  facet_grid(Sex ~ Month)
```

####Part 2: Habitat parameters for Aegla crabs
```{r}
file_path <- "~/bi485/beardslee_jordan/data/aegla_crabs.csv"

crabs <- aegla_crabs
crabs <- subset(crabs, select = -c(Site,AT,WT,Shrimp,Aeglam,AeglaFe,AeglaFo))
crabs <- drop_na(crabs)
crabs <- crabs %>%
  mutate(row_order = 1:nrow(.))
View(crabs)
```

####Explore the data
```{r}
p_width <- crabs %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = Width)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
```

```{r}
p_Depth <- crabs %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = Depth)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
```

```{r}
p_flow <- crabs %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = Flow)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
```

```{r}
p_pH <- crabs %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = pH)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
```

```{r}
p_TDS <- crabs %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = TDS)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
```

```{r}
p_cond <- crabs %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = Cond)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
```

```{r}
p_N <- crabs %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = N)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
```


```{r}
p_Ni <- crabs %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = Ni)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
```

```{r}
p_Nt <- crabs %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = Nt)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
```

```{r}
p_phos <- crabs %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = Phos)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
```

```{r}
p_Mg <- crabs %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = Mg)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
```

```{r}
p_Ca <- crabs %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = Ca)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
```

```{r}
p_Crab <- crabs %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = Crab)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
```

```{r}
p_width + p_Depth + p_flow + p_pH + p_TDS + p_cond + p_N + p_Ni + p_Nt + p_phos + p_Mg + p_Ca + p_Crab + plot_layout(ncol = 6)
```
Question 1: Crab, Depth, and Ph have the more obvious outliers. These three have data points that are most obviously broken away from the main group. Crabs has two that don't remain within reasonable devation from the main data, Depth has one point well away from the other, ver closely packed values, and PH has a point that is nowhere near high enough to be normal.

```{r}
Crab_plot <- crabs %>% 
  ggplot(aes(x = row_order, 
             y = Crab)) +
  geom_boxplot(width = 0.25) 
Crab_plot
```

```{r}
ph_plot <- crabs %>% 
  ggplot(aes(x = row_order, 
             y = pH)) +
  geom_boxplot(width = 0.25) 
ph_plot
```

```{r}
depth_plot <- crabs %>% 
  ggplot(aes(x = row_order, 
             y = Depth)) +
  geom_boxplot(width = 0.25) 
depth_plot
```
By the looks of it, Crabs has the most outliers since there are several points completely outside the error bars.


```{r}
plot_width <- crabs %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = Width)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
```

```{r}
plot_depth <- crabs %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = Depth)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
```

```{r}
plot_flow <- crabs %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = Flow)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
```

```{r}
plot_crabs <- crabs %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = Crab)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
```

```{r}
plot_width + plot_depth + plot_flow + plot_crab + plot_layout(ncol = 2)
```

####Histograms

```{r}
Crabsh1 <- crabs %>%
  ggplot() +
  geom_histogram(aes(x = Mg),
                 binwidth = 1,
                 boundary = 0,
                 closed = "right",
                 color = "black") +
  scale_x_continuous(breaks = seq(2, 82, by = 4))
Crabsh1
```

```{r}
Crabsh2 <- crabs %>%
  ggplot() +
  geom_histogram(aes(x = TDS),
                 binwidth = 3,
                 boundary = 0,
                 closed = "right",
                 color = "black") +
  scale_x_continuous(breaks = seq(2, 82, by = 4))
Crabsh2
```

```{r}
Crabsh3 <- crabs %>%
  ggplot() +
  geom_histogram(aes(x = Cond),
                 binwidth = 3,
                 boundary = 0,
                 closed = "right",
                 color = "black") +
  scale_x_continuous(breaks = seq(2, 82, by = 4))
Crabsh3
```
####Density Plots

```{r}
Crabsd1 <- crabs %>% 
  ggplot() +
  geom_density(aes(x = Mg)) +
  scale_x_continuous(breaks = seq(0,40, by = 5))
Crabsd1
```

```{r}
Crabsd2 <- crabs %>% 
  ggplot() +
  geom_density(aes(x = TDS)) +
  scale_x_continuous(breaks = seq(0,40, by = 5))
Crabsd2
```

```{r}
Crabsd3 <- crabs %>% 
  ggplot() +
  geom_density(aes(x = Cond)) +
  scale_x_continuous(breaks = seq(0,40, by = 5))
Crabsd3
```

####Patchwork
```{r}
histograms <- Crabsh1 + Crabsh2 + Crabsh3 + plot_layout(nrow = 3)

histograms
```

```{r}
density <- Crabsd1 + Crabsd2 + Crabsd3 + plot_layout(nrow = 3)

density
```

####Explore Relationship Between Variables
```{r}
crabs %>% ggpairs(collumns = 1|14)
```

