---
title: "HW 06 Part 1: Complete the sections"
author: "Jordan Beardslee"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

***
####Part 2 Darters

```{r}
library(tidyverse)

library(knitr)

library(smatr)
```

####Import and Wrangle
```{r}
file_path <- "~/bi485/beardslee_jordan/data/darter.txt"
widths <- fwf_widths(c(3, 11, 1, 1, 1, 3, 4, 4, 4, 3, 3, 2, 2))
darter <- read_fwf("darters.txt", comment = "#", col_positions = widths)
col_names <- c("id", "species", "riffle", "age", "sex", "length", "depth", "velocity", "temperature", "majsub", "minsub", "majtype", "mintype")
colnames(darter) <- col_names
darter <- filter(darter, species == "spectiabile" | species == "caeruleum" | species == "blennioides")
```
```{r}
mutate(darter,
       total_substrate = majsub + minsub)
```
```{r}
mutate(darter,
       riffle = ifelse(riffle == 1,
                       "riffle_1",
                       "riffle_2"))
```
```{r}
mutate(darter,length = length * 10)
```
```{r}
mutate(darter, sex = ifelse(sex == "f", "Female", "Male"))
```
####Summarise data

```{r}
darter <- group_by(darter, species)
darters_summary <- darter %>%
  summarise(length <- mean(length, na.rm = TRUE),
                          depth = mean(depth, na.rm = TRUE),
                                       velocity = mean(velocity, na.rm= TRUE))

darters_summary
```
####Plot<1.>
```{r}
ggplot(data = darter, mapping = aes(x = species, y = length)) +
      geom_point(data = darter, mapping = aes(x = species, y = length), shape = 10) +
      coord_flip() +
      facet_wrap(~ riffle, nrow = 2) +
      xlab("species") +
      ylab("Length (mm)")


```
Riffle 2 means seem lower than riffle 1, caeruleum seems about even on both tables, while blennioides seems to be slightly higher in riffle 2.
####Plot<2.>
```{r}
ggplot(data = darter, mapping = aes(x = species, y = length)) +
  stat_summary(
    fun.ymin = min,
    fun.ymax = max,
    fun.y = median,
    geom = "point",
    size = 2.2) +
  stat_summary(
    fun.y = mean,
    size = 3,
    color = "maroon"
  ) +
      geom_point(data = darter, fill = "grey70", mapping = aes(x = species, y = length)) +
      coord_flip() +
      facet_wrap(~ riffle, nrow = 2) +
      xlab("species") +
      ylab("Length (mm)")
```

####Stress and corticosterones in birds
```{r}
?locale
file_path <- "~/bi485/beardslee_jordan/data/darter.txt"

feathers <- read_csv("feathers.csv")

locale(decimal_mark = ".", grouping_mark = ",")

feathers
```
Data appears to be tidy, each variable is in a column and capture dates are aligned in rows. 

```{r}
names(feathers) <- c("capture_date", "day", "site", "urban_score", "cortF", "cortB", "sex", "smi", "tarsus", "mass", "basal_cort", "stress_cort")

feathers

```
```{r}
mean_tarsus <- summarise(feathers, mean_tarsus = mean(tarsus, na.rm = TRUE))
```
```{r}
mean_tarsus
```
```{r}
feathers
```

```{r}
feathers <- feathers %>%
  mutate(log_mass = log(mass),
         log_tarsus= log(tarsus))
```

```{r}
log_mass
```
```{r}
log_tarsus
```

```{r}
major_axis <- sma(log_mass ~ log_tarsus, data = feathers)

```
```{r}
summary(major_axis)

```
```{r}
coef(major_axis)

b <- coef(major_axis)
```

```{r}
new_smi <- (feathers$mass * ((mean_tarsus/feathers$tarsus)^b))


```
The values are basically identical in both elevation and slope.

```{r}
group_by(feathers, site)
```

```{r}
mean_basal_cort <- mean(feathers$basal_cort)

mean_stress_cort <- mean(feathers$stress_cort)

sd_basal_cort <- sd(feathers$basal_cort, na.rm = TRUE)

sd_stress_cort <- sd(feathers$stress_cort, na.rm = TRUE)

se_basal_cort <- sd_basal_cort/(sqrt(43))

se_stress_cort <- sd_stress_cort/(sqrt(43))

se_basal_cort

se_stress_cort
```


```{r}
new_smi
```



```{r}
feathers
```


####Plot<1.>
```{r}
ggplot(data = feathers, mapping = aes(x = site, y = "new_smi")) + 
  geom_boxplot()
```


####Plot<2.>

```{r}
basal_min <- mean_basal_cort - sd_basal_cort

basal_max <- mean_basal_cort + sd_basal_cort
```


```{r}
ggplot(data = feathers, mapping = aes(x = site, y = basal_cort)) + 
  geom_point() +
  geom_errorbar(ymin = basal_min, ymax = basal_max)
```
####Plot<3.>
```{r}
max_stress <- mean_stress_cort + sd_stress_cort

min_stress <- mean_stress_cort - sd_stress_cort
```


```{r}
ggplot(data = feathers, mapping = aes(x = stress_cort, y = site)) + 
  geom_point() +
  geom_linerange(ymin = min_stress, ymax = max_stress)
```

